---
title: "Calculations and Figures using Continuous Dataset"
output:
  html_document:
    toc: true
    df_print: paged
authors: Kevin Zhang
---


<!-- 1. Set up continuous dataset (Line 31) -->
<!-- 2. Comparison of training and testing datasets (Line 78) -->
<!-- 3. Find missing % at listing (Line 267) -->
<!-- 4. Impute values into unblocked dataset (Line 295) -->
<!-- 5. Cox PH with US-CRS predictors on unblocked and generalized c-index (Line 407) -->


```{r, include=F, warning=F, echo=F, comment=NA}
rm(list=ls()); invisible(gc())
if (!require('pacman')) {install.packages('pacman')}
library(pacman)
pacman::p_load(tidyverse, splines, knitr, kableExtra, glmnet, caret, gtsummary, 
               pROC, survival, DescTools, rsample)
knitr::opts_knit$set(root.dir = 'C:/Users/Kevin (Work)/Desktop/US-CRS')
```


---------------------------------


# 1. Set up continuous dataset
Continuous dataset is obtained from Section 5 of *6week_death_data_prep.Rmd*. 


```{r, echo=F, warning=F, comment=NA}
load('./intervals_v17_continuous.RData')


## Keep only relevant variables
df <- df %>%
  select(PX_ID, CenterId, t_start, t_stop, last_group, status,
         systolicBP, PCWP, central_venous_pressure, cardiac_output, 
         AST, creatinine, bilirubin, albumin, sodium, BNP, BNP_NT_Pro,
         heart_rate, diastolicBP, PASP, PADP, hemoglobin, dialysis, peripheral_vascular, 
         status, ECMO, BiVAD_no_discharge, temp_surg_LVAD, IABP, ECMO, 
         death_flag, age, status_initial, LVAD, BiVAD, transplant, perc_LVAD)


## Add a train-test indicator
load('./train_ids.RData')
df$train_test <- 'Test'
df$train_test[df$CenterId %in% center_ids_train] <- 'Train'


## General death; death in final status change
df <- df %>%
  group_by(PX_ID) %>%
  mutate(death = case_when(
    death_flag == '1' & transplant == '0' & last_group == '1' ~ 1,
    TRUE ~ 0)) %>%
  select(-c(death_flag, transplant)) %>%
  
  mutate(death=max(death))


## Certain variables need to be obtained from cand_thor since demographics not used in model
df_cand_thor <- haven::read_sas('./cand_thor.sas7bdat')
df_cand_thor <- df_cand_thor %>% select(PX_ID, CAN_GENDER, CAN_RACE, CAN_HGT_CM, CAN_WGT_KG) %>% unique()
df <- merge(df, df_cand_thor, by='PX_ID', all.x=T, all.y=F)
df$bsa <- 0.007184 * (df$CAN_HGT_CM ^ 0.425) * (df$CAN_WGT_KG ^ 0.725)
rm(df_cand_thor)
```


---------------------------------


# 2. Comparison of training and testing datasets

Includes these initial variables at listing:

- 6 Status at Listing
- Age at Listing
- Sex
- Race
- Albumin
- Bilirubin
- eGFR
- Sodium
- BNP
- BNP Type
- Pulmonary Systolic BP, Pulmonary Diastolic BP
- Central Venous Pressure
- CPO
- API
- MCS Components (ECMO, temporary surgical LVAD, BiVAD without discharge)
- IABP
- Durable LVAD
- Percutaneous VAD


```{r, warning=F, comment=NA, fig.align='center'}
df_table1 <- df %>%
  select(PX_ID, train_test, status, age, albumin,
         bilirubin, creatinine, sodium, BNP, BNP_NT_Pro,
         PASP, PADP, central_venous_pressure,
         systolicBP, diastolicBP, cardiac_output, 
         CAN_GENDER, CAN_RACE, dialysis,
         ECMO, temp_surg_LVAD, BiVAD_no_discharge,
         IABP, LVAD, perc_LVAD) %>%
  group_by(PX_ID) %>%
  filter(row_number() == 1) %>%
  mutate(age = as.numeric(age))


### Recreate eGFR and race
df_table1 <- df_table1 %>%
  mutate(eGFR = case_when(
    CAN_GENDER == 'F' & !is.na(creatinine) ~ 142 * (pmin((creatinine / 0.7), 1)^(-0.241)) *
      (pmax((creatinine / 0.7), 1)^(-1.2)) * 0.9938^(age) * 1.012,
    
    CAN_GENDER == 'M' & !is.na(creatinine) ~ 142 * (pmin((creatinine / 0.9), 1)^(-0.302)) *
      (pmax((creatinine / 0.7), 1)^(-1.2)) * 0.9938^(age))) %>%
  
  mutate(race = case_when(
    CAN_RACE == '8' ~ 'White',
    CAN_RACE == '16' ~ 'Black or AA',
    CAN_RACE == '32' ~ 'American Indian or Alaska Native',
    CAN_RACE == '64' ~ 'Asian',
    CAN_RACE == '128' ~ 'Native Hawaiian or Other Pacific Islander',
    CAN_RACE == '256' ~ 'Arab or Middle Eastern',
    CAN_RACE == '512' ~ 'Indian Sub-continent',
    CAN_RACE == '1024' ~ 'Unknown',
    CAN_RACE == '2000' ~ 'Hispanic or Latino',
    TRUE ~ 'Multi-Racial')) 


### Recreate CPO and API
df_table1 <- df_table1 %>%   
  mutate(
    cpo = (1/541) * cardiac_output * (((2/3)*systolicBP) + ((1/3)*diastolicBP)),
    papi = (PASP - PADP) / central_venous_pressure)

df_table1$eGFR[df_table1$dialysis == '1'] <- 0
df_table1$papi[df_table1$central_venous_pressure == 0] <- 1
df_table1$papi[df_table1$papi < 1 & !is.na(df_table1$papi)] <- 1



df_table1$train_test <- factor(df_table1$train_test, levels = c('Train', 'Test'))
df_table1 <- df_table1[ ,c('train_test', 'age', 'CAN_GENDER', 'race', 
                           'status', 'albumin', 'bilirubin', 'eGFR',
                           'sodium', 'BNP', 'BNP_NT_Pro', 'cardiac_output', 
                           'PASP', 'PADP', 'systolicBP', 'diastolicBP',
                           'central_venous_pressure', 'cpo', 'papi',
                           'ECMO', 'temp_surg_LVAD', 'BiVAD_no_discharge',
                           'IABP', 'LVAD', 'perc_LVAD')]
df_table1$BNP_NT_Pro <- as.factor(df_table1$BNP_NT_Pro)
df_table1$ECMO <- as.factor(df_table1$ECMO)
df_table1$temp_surg_LVAD <- as.factor(df_table1$temp_surg_LVAD)
df_table1$BiVAD_no_discharge <- as.factor(df_table1$BiVAD_no_discharge)
df_table1$IABP <- as.factor(df_table1$IABP)
df_table1$LVAD <- as.factor(df_table1$LVAD)
df_table1$perc_LVAD <- as.factor(df_table1$perc_LVAD)


df_table1 %>%
  ungroup() %>%
  tbl_summary(by = train_test, 
              missing = 'no', 
              statistic = all_continuous() ~ "{mean} ({sd})",
              label = c(age ~ 'Age at Listing, mean (SD), y',
                        CAN_GENDER ~ 'Gender, No. (%)',
                        race ~ 'Race, No. (%)',
                        status ~ '6-Status, No. (%)',
                        albumin ~ 'Albumin, mean (SD), g/dL',
                        bilirubin ~ 'Bilirubin, mean (SD), mg/dL',
                        eGFR ~ 'eGFR, mean (SD), mL/min/1.73m2',
                        sodium ~ 'Sodium, mean (SD), mEq/L',
                        BNP ~ 'BNP, mean (SD), pg/mL',
                        BNP_NT_Pro ~ 'NT pro-BNP, No. (%)',
                        cardiac_output ~ 'Cardiac output, mean (SD), L/min',
                        PASP ~ 'Pulmonary Systolic BP, mean (SD), mm Hg',
                        PADP ~ 'Pulmonary Diastolic BP, mean (SD), mm Hg',
                        systolicBP ~ 'Systolic BP, mean (SD), mm Hg',
                        diastolicBP ~ 'Diastolic BP, mean (SD), mm Hg',
                        central_venous_pressure ~ 'Central Venous Pressure, mean (SD), mm Hg',
                        cpo ~ 'Cardiac Power Output, mean (SD), watts',
                        papi ~ 'Pulmonary Artery Pulsatility Index, mean (SD)',
                        ECMO ~ 'ECMO, No. (%)',
                        temp_surg_LVAD ~ 'Temporary Surgical LVAD, No. (%)',
                        BiVAD_no_discharge ~ 'BiVAD, No Discharge, No. (%)',
                        IABP ~ 'IABP, No. (%)',
                        LVAD ~ 'Durable LVAD, No. (%)',
                        perc_LVAD ~ 'Percutaneous LVAD, No. (%)')) %>%
  modify_header(label = "**Variable**") %>%
  bold_labels()

df_table1 %>%
  ungroup() %>%
  tbl_summary(by = train_test, 
              missing = 'no', 
              label = c(age ~ 'Age at Listing, mean (SD), y',
                        CAN_GENDER ~ 'Gender, No. (%)',
                        race ~ 'Race, No. (%)',
                        status ~ '6-Status, No. (%)',
                        albumin ~ 'Albumin, mean (SD), g/dL',
                        bilirubin ~ 'Bilirubin, mean (SD), mg/dL',
                        eGFR ~ 'eGFR, mean (SD), mL/min/1.73m2',
                        sodium ~ 'Sodium, mean (SD), mEq/L',
                        BNP ~ 'BNP, mean (SD), pg/mL',
                        BNP_NT_Pro ~ 'NT pro-BNP, No. (%)',
                        cardiac_output ~ 'Cardiac output, mean (SD), L/min',
                        PASP ~ 'Pulmonary Systolic BP, mean (SD), mm Hg',
                        PADP ~ 'Pulmonary Diastolic BP, mean (SD), mm Hg',
                        systolicBP ~ 'Systolic BP, mean (SD), mm Hg',
                        diastolicBP ~ 'Diastolic BP, mean (SD), mm Hg',
                        central_venous_pressure ~ 'Central Venous Pressure, mean (SD), mm Hg',
                        cpo ~ 'Cardiac Power Output, mean (SD), watts',
                        papi ~ 'Pulmonary Artery Pulsatility Index, mean (SD)',
                        ECMO ~ 'ECMO, No. (%)',
                        temp_surg_LVAD ~ 'Temporary Surgical LVAD, No. (%)',
                        BiVAD_no_discharge ~ 'BiVAD, No Discharge, No. (%)',
                        IABP ~ 'IABP, No. (%)',
                        LVAD ~ 'Durable LVAD, No. (%)',
                        perc_LVAD ~ 'Percutaneous LVAD, No. (%)')) %>%
  modify_header(label = "**Variable**") %>%
  bold_labels()

df_table1 %>%
  ungroup() %>%
  tbl_summary(by = train_test, 
              missing = 'ifany', 
              statistic = all_continuous() ~ "{mean} ({sd})",
              label = c(age ~ 'Age at Listing, mean (SD), y',
                        CAN_GENDER ~ 'Gender, No. (%)',
                        race ~ 'Race, No. (%)',
                        status ~ '6-Status, No. (%)',
                        albumin ~ 'Albumin, mean (SD), g/dL',
                        bilirubin ~ 'Bilirubin, mean (SD), mg/dL',
                        eGFR ~ 'eGFR, mean (SD), mL/min/1.73m2',
                        sodium ~ 'Sodium, mean (SD), mEq/L',
                        BNP ~ 'BNP, mean (SD), pg/mL',
                        BNP_NT_Pro ~ 'NT pro-BNP, No. (%)',
                        cardiac_output ~ 'Cardiac output, mean (SD), L/min',
                        PASP ~ 'Pulmonary Systolic BP, mean (SD), mm Hg',
                        PADP ~ 'Pulmonary Diastolic BP, mean (SD), mm Hg',
                        systolicBP ~ 'Systolic BP, mean (SD), mm Hg',
                        diastolicBP ~ 'Diastolic BP, mean (SD), mm Hg',
                        central_venous_pressure ~ 'Central Venous Pressure, mean (SD), mm Hg',
                        cpo ~ 'Cardiac Power Output, mean (SD), watts',
                        papi ~ 'Pulmonary Artery Pulsatility Index, mean (SD)',
                        ECMO ~ 'ECMO, No. (%)',
                        temp_surg_LVAD ~ 'Temporary Surgical LVAD, No. (%)',
                        BiVAD_no_discharge ~ 'BiVAD, No Discharge, No. (%)',
                        IABP ~ 'IABP, No. (%)',
                        LVAD ~ 'Durable LVAD, No. (%)',
                        perc_LVAD ~ 'Percutaneous LVAD, No. (%)')) %>%
  modify_header(label = "**Variable**") %>%
  bold_labels()
```


---------------------------------


# 3. Find missing % at listing

```{r, warning=F, comment=NA, fig.align='center'}
df_missings <- df %>%
  select(PX_ID, BNP, sodium, bilirubin, albumin, cardiac_output,
         central_venous_pressure, PASP, PADP, 
         systolicBP, diastolicBP, PCWP,
         CAN_GENDER, creatinine, age) %>%
  group_by(PX_ID) %>%
  filter(row_number() == 1) %>%
  mutate(age = as.numeric(age))


df_missings <- df_missings %>%
  mutate(eGFR = case_when(
    CAN_GENDER == 'F' & !is.na(creatinine) ~ 142 * (pmin((creatinine / 0.7), 1)^(-0.241)) *
      (pmax((creatinine / 0.7), 1)^(-1.2)) * 0.9938^(age) * 1.012,
    
    CAN_GENDER == 'M' & !is.na(creatinine) ~ 142 * (pmin((creatinine / 0.9), 1)^(-0.302)) *
      (pmax((creatinine / 0.7), 1)^(-1.2)) * 0.9938^(age)))

100*colMeans(is.na(df_missings))
```


---------------------------------


# 4. Impute values into unblocked dataset

### Impute using the same rules

```{r, warning=F, comment=NA}
df <- df %>%
  group_by(PX_ID) %>%
  fill(status, .direction = 'down')

df$hemoglobin[is.na(df$hemoglobin) & df$CAN_GENDER == 'M'] <- 14.9
df$hemoglobin[is.na(df$hemoglobin) & df$CAN_GENDER == 'F'] <- 13.3

df$bilirubin[is.na(df$bilirubin)] <- median(df$bilirubin, na.rm=T)
df$albumin[is.na(df$albumin)] <- median(df$albumin, na.rm=T)

df$BNP[is.na(df$BNP) & df$BNP_NT_Pro == '0'] <- 
  median(df$BNP[df$BNP_NT_Pro == '0'], na.rm=T)
df$BNP[is.na(df$BNP) & df$BNP_NT_Pro == '1'] <- 
  median(df$BNP[df$BNP_NT_Pro == '1'], na.rm=T)

df <- df %>%
  mutate(cardiac_output = ifelse((is.na(cardiac_output) & LVAD == '1'),  2.5*bsa, cardiac_output)) %>%
  mutate(cardiac_output = ifelse((is.na(cardiac_output) & LVAD != '1'),  2.2*bsa, cardiac_output)) %>%
  select(-c(bsa))


df$central_venous_pressure[is.na(df$central_venous_pressure) & df$LVAD == '1'] <- 10
df$PASP[is.na(df$PASP) & df$LVAD == '1'] <- 35
df$PADP[is.na(df$PADP) & df$LVAD == '1'] <- 15
df$PCWP[is.na(df$PCWP) & df$LVAD == '1'] <- 15
df$systolicBP[is.na(df$systolicBP) & df$LVAD == '1'] <- 100
df$diastolicBP[is.na(df$diastolicBP) & df$LVAD == '1'] <- 80
df$heart_rate[is.na(df$heart_rate) & df$LVAD == '1'] <- 8.5


df$central_venous_pressure[is.na(df$central_venous_pressure) & df$LVAD != '1'] <- 12
df$PASP[is.na(df$PASP) & df$LVAD != '1'] <- 38
df$PADP[is.na(df$PADP) & df$LVAD != '1'] <- 18
df$PCWP[is.na(df$PCWP) & df$LVAD != '1'] <- 18
df$systolicBP[is.na(df$systolicBP) & df$LVAD != '1'] <- 110
df$diastolicBP[is.na(df$diastolicBP) & df$LVAD != '1'] <- 80
df$heart_rate[is.na(df$heart_rate) & df$LVAD != '1'] <- 8.5
```


### Recreate short MCS ever, eGFR, cpo, api

```{r, warning=F, comment=NA, fig.align='center'}
df <- df %>%
  mutate(ECMO_ever = case_when(ECMO == '1' ~ '1', TRUE ~ NA_character_),
         temp_surg_ever = case_when(temp_surg_LVAD == '1' ~ '1', TRUE ~ NA_character_),
         BiVAD_no_discharge_ever = case_when(BiVAD_no_discharge == '1' ~ '1', TRUE ~ NA_character_)) %>%
  
  fill(ECMO_ever, .direction = 'down') %>%
  fill(temp_surg_ever, .direction = 'down') %>%
  fill(BiVAD_no_discharge_ever, .direction = 'down') %>%
  
  mutate(ECMO_ever = case_when(is.na(ECMO_ever) ~ '0', TRUE ~ ECMO_ever)) %>%
  mutate(temp_surg_ever = case_when(is.na(temp_surg_ever) ~ '0', TRUE ~ temp_surg_ever)) %>%
  mutate(BiVAD_no_discharge_ever = case_when(is.na(BiVAD_no_discharge_ever) ~ '0', TRUE ~ BiVAD_no_discharge_ever)) %>%
  
  
  mutate(short_MCS_French = 
           case_when(
             (ECMO == '1' & status == '1') | IABP == '1' ~ 1,
             TRUE ~ 0)) %>% 
  mutate(short_MCS_French_new = 
           case_when((ECMO == '1' & status == '1') ~ 1,
             TRUE ~ 0)) %>% 
  mutate(short_MCS_ever = 
           case_when(
             ECMO_ever == '1' | temp_surg_ever == '1' | BiVAD_no_discharge_ever == '1' ~ 1,
             TRUE ~ 0)) %>%
  
 mutate(eGFR = case_when(
    
    CAN_GENDER == 'F' & !is.na(creatinine) ~ 142 * (pmin((creatinine / 0.7), 1)^(-0.241)) *
      (pmax((creatinine / 0.7), 1)^(-1.2)) * 0.9938^age * 1.012,
    
    CAN_GENDER == 'M' & !is.na(creatinine) ~ 142 * (pmin((creatinine / 0.9), 1)^(-0.302)) *
      (pmax((creatinine / 0.7), 1)^(-1.2)) * 0.9938^age)) %>%
  
  mutate(
    cpo = (1/541) * cardiac_output * (((2/3)*systolicBP) + ((1/3)*diastolicBP)),
    papi = (PASP - PADP) / central_venous_pressure)


df$eGFR[df$dialysis == '1'] <- 0
df$eGFR[is.na(df$eGFR)] <- 100

df$papi[df$central_venous_pressure == 0] <- 1
df$papi[df$papi < 1 & !is.na(df$papi)] <- 1

df$LVAD <- as.factor(df$LVAD)


## Remove 0-time blocks
df <- subset(df, df$t_start != df$t_stop) 
df$last_group <- NULL


df <- df %>%
  group_by(PX_ID) %>%
  mutate(death = case_when(
    death == '1' & row_number() == n() ~ 1,
    TRUE ~ 0))
```


---------------------------------


# 5. Cox PH with US-CRS predictors on unblocked and generalized c-index

```{r, warning=F, message=F, comment=NA}
load('./model_frenchCRS_new.RData')
load('./model_USCRS.RData')

pacman::p_load(risksetROC)
df$BNP_NT_Pro <- as.factor(df$BNP_NT_Pro)

## Basic Cox model
model_us <- coxph(Surv(t_start, t_stop, death) ~ albumin + log(bilirubin + 1) + 
        eGFR + sodium + (log(BNP + 1):BNP_NT_Pro) + LVAD + short_MCS_ever, 
      data = df[df$train_test == 'Train',])

df$us_predict <- predict(model_final, df)
df$french_predict <- predict(model_french_new, df)


df$status_ranking <- 10 - as.numeric(df$status_initial)
df <- df %>% 
  filter(train_test == 'Test') %>%
  filter(!is.na(status_ranking) & !is.na(us_predict) & !is.na(french_predict))


### Functions to get generalized c-index
get_overall_six <- function(data){
  r <- risksetAUC(
    Stime= data$t_stop,
    entry = data$t_start,
    status = data$death,
    marker = data$status_ranking,
    method = 'Cox',
    tmax = 365*2,
    plot = F)
  return(r$Cindex)
}

get_overall_french <- function(data){
  r <- risksetAUC(
    Stime= data$t_stop,
    entry = data$t_start,
    status = data$death,
    marker = data$french_predict,
    method = 'Cox',    
    tmax = 365*2,
    plot = F)
  return(r$Cindex)
}

get_overall_us <- function(data){
  r <- risksetAUC(
    Stime= data$t_stop,
    entry = data$t_start,
    status = data$death,
    marker = data$us_predict,
    method = 'Cox',
    tmax = 365*2,
    plot = F)
  return(r$Cindex)
}

df_short <- df[ ,c('PX_ID', 't_start', 't_stop', 'death', 'us_predict', 
                  'french_predict', 'status_ranking')]
df_nest <- df_short %>% nest(-PX_ID)
```


```{r, warning=F, message=F, comment=NA}
set.seed(200)
get_overall_six(df)

n_iterations <- 2000
bs <- rsample::bootstraps(df_nest, times = n_iterations)
auc_list_six <- vector(length = n_iterations)


for (i in 1:n_iterations) {
  result_df <- as.tibble(bs$splits[[i]]) %>% unnest() %>% as.data.frame()
  auc_list_six[i] <- get_overall_six(result_df)
}


quantile(auc_list_six, probs = c(0.025, 0.975))
```

```{r, warning=F, message=F, comment=NA}
set.seed(200)
get_overall_french(df)

auc_list_french <- vector(length = n_iterations)

for (i in 1:n_iterations) {
  result_df <- as.tibble(bs$splits[[i]]) %>% unnest() %>% as.data.frame()
  auc_list_french[i] <- get_overall_french(result_df)
}

quantile(auc_list_french, probs = c(0.025, 0.975))
```

```{r, warning=F, message=F, comment=NA}
set.seed(200)
get_overall_us(df)

auc_list_us <- vector(length = n_iterations)

for (i in 1:n_iterations) {
  result_df <- as.tibble(bs$splits[[i]]) %>% unnest() %>% as.data.frame()
  auc_list_us[i] <- get_overall_us(result_df)
}

quantile(auc_list_us, probs = c(0.025, 0.975))
```
